\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2016-12-21 mSystems/rev/mSys_MAIN_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

% release codenumber (update it here!)
\newcommand\releasecode{final}

%%% PACKAGES
\usepackage{lmodern} % Activa fuentes y macros de Latin Modern
\usepackage[T1]{fontenc} % set font encoding
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage[english]{babel} 
\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
%%\usepackage{calligra}
\usepackage[printwatermark]{xwatermark}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{parskip}
%%\usepackage{float}
\usepackage[nolists, nomarkers]{endfloat}
\usepackage{moreverb} % for verbatim output

@@

% watermarks (time consuming!)
\newwatermark[allpages,color=PineGreen!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
\newwatermark[allpages,color=PineGreen!12,angle=0,scale=1,xpos=0,ypos=137]{
\emph{mSystems} \LaTeX\ manuscript \releasecode\ submission}

@@

% word count (Need --enable-write18 or --shell-escape)
\immediate\write18{texcount -opt=option.tc \jobname.tex > wordcount.aux} 
\newcommand\wordcount{\verbatiminput{wordcount.aux}}

@@

\author[1,2,$*$]{Jose Manuel Martí}
\author[1,2,3,$*$]{Daniel Martínez-Martínez}
\author[2]{Manuel Peña}
\author[1,2]{César Gracia}
\author[1,3,4,5]{Amparo Latorre}
\author[1,3,4,5,\#]{Andrés Moya}
\author[1,2,\#]{Carlos P. Garay}

@@

%TC:newcounter abswords Words in abstract
%TC:macro \abstract [abswords]
{\abstract{Animal microbiota (human included) plays an important role keeping healthy the physiological status of the host. Increasing research activity is dedicated to understand how changes in composition and function of the microbiota are associated to disease or not. We analyze 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbial composition reveal significant differences due to factors such us dietary changes, antibiotic intake, age or disease. Here we show that a fluctuation scaling law describes the temporal changes in the gut microbiota. This law allows to estimate the temporal variability of the microbial population and quantitatively characterizes the path toward disease by a noise-induced phase transition. The estimation of the systemic parameters for follow-up studies may have clinical use and, more generally, applications in other fields where it is important to know if a given community is stable or not.}}

%TC:newcounter impwords Words in importance section
%TC:macro \importance [impwords]
\importance{Human microbiota is tightly associated to the health status of a person. Here we analyse the microbial composition of several subjects under different conditions, over a time span that ranges from days to months. Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas. This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system.}

@@

%===================================== SECTIONS =======================================
%%% Introduction
\section*{Introduction}
\input{mSys_intro.tex}
%\clearpage
%%% Result
\section*{Results}
\input{mSys_results.tex}
%\clearpage
%%% Discussion
\section*{Discussion}
\input{mSys_discussion.tex}
%\clearpage    
%%% Material & Methods
\section*{Materials and Methods}
\input{mSys_MandM.tex}   
%\clearpage

@@

%% Discussion
\bibitem {kilpatrick} {\bf Kilpatrick a M, Ives a R.} 2003. Species interactions can explain Taylor’s power law for ecological time series. Nature {\bf 422}:65–68.
\bibitem {ballantyne} {\bf Ballantyne IV F, J. Kerkhoff A.} 2007. The observed range for temporal mean-variance scaling exponents can be explained by reproductive correlation. Oikos {\bf 116}:174–180.
\bibitem {joao} {\bf Stein RR, Bucci V, Toussaint NC, Buffie CG, Rätsch G, Pamer EG, Sander C, Xavier JB.}  2013. Ecological modeling from time-series inference: insight into dynamics and stability of intestinal microbiota. PLoS Comput Biol {\bf 9}:e1003388.
\bibitem {mehta} {\bf Fisher CK, Mehta P.} 2014. Identifying keystone species in the human gut microbiome from metagenomic timeseries using sparse linear regression. PLoS One {\bf 9}:e102451.
\bibitem {bucci}  {\bf Bucci V, Tzen B, Li N, Simmons M, Tanoue T, Bogart E, Deng L, Yeliseyev V, Delaney ML, Liu Q, Olle B, Stein RR, Honda K, Bry L, Gerber GK.} 2016. MDSINE: Microbial Dynamical Systems INference Engine for microbiome time-series analyses. Genome Biol {\bf 17}:121.
\bibitem {koenig} {\bf Koenig JE, Spor A, Scalfone N, Fricker AD, Stombaugh J, Knight R, Angenent LT, Ley RE.} 2011. Succession of microbial consortia in the developing infant gut microbiome. Proc Natl Acad Sci {\bf 108}:4578–4585.
\bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.

>>>>

% release codenumber (update it here!)
\newcommand\releasecode{REVISION draft A}

%%% PACKAGES
\usepackage{lmodern} % Activa fuentes y macros de Latin Modern
\usepackage[T1]{fontenc} % set font encoding
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage[english]{babel} 
\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
%%\usepackage{calligra}
\usepackage[printwatermark]{xwatermark}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{parskip}
\usepackage{moreverb} % for verbatim output

@@

% watermarks (time consuming!)
\newwatermark[allpages,color=Mulberry!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
\newwatermark[allpages,color=Mulberry!12,angle=0,scale=1,xpos=0,ypos=137]{
\emph{mSystems} \LaTeX\ manuscript \releasecode}

@@

% word count (Need --enable-write18 or --shell-escape)
\immediate\write18{texcount -opt=option_rev.tc \jobname.tex > wordcount_rev.aux} 
\newcommand\wordcount{\verbatiminput{wordcount_rev.aux}}

% Floating stuff
\usepackage[nolists, nomarkers]{endfloat}
%\renewcommand{\processdelayedfloats}{}  % Avoid float processing
\usepackage{newfloat}
\DeclareFloatingEnvironment[
	fileext=lsf,
	listname={List of Supplementary Figures},
	name=Supplementary Figure,
	placement=p
]{supfig}
\DeclareDelayedFloat{supfig}{Supplementary Figures}

@@

\author[1,2,$*$]{Jose Manuel Martí}
\author[1,2,3,$*$]{Daniel Martínez-Martínez}
\author[2]{Manuel Peña}
\author[1,2]{César Gracia}
\author[1]{Teresa Rubio}
\author[1,3,4,5]{Amparo Latorre}
\author[1,3,4,5,\#]{Andrés Moya}
\author[1,2,\#]{Carlos P. Garay}

@@

%% TeXcount commands (in addition to option_rev.tc)
%TC:newcounter supfig Number of Supplemental Figures
%TC:macrocount beginsupfig [supfig]
%TC:newcounter supfigwords Words Supplemental Figures
%TC:envir supfig [] supfigwords
%TC:newcounter abswords Words in abstract
%TC:macro \abstract [abswords]
%TC:newcounter impwords Words in importance section
%TC:macro \importance [impwords]

{\abstract{Animal microbiota (human included) plays an important role keeping healthy the physiological status of the host. Increasing research activity is dedicated to understand how changes in composition and function of the microbiota are associated to disease or not. We analyze 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbial composition reveal significant differences due to factors such us dietary changes, antibiotic intake, age or disease. Here we show that a fluctuation scaling law describes the temporal changes in the gut microbiota. This law allows to estimate the temporal variability of the microbial population and quantitatively characterizes the path toward disease by a noise-induced phase transition. The estimation of the systemic parameters for follow-up studies may have clinical use and, more generally, applications in other fields where it is important to know if a given community is stable or not.}}

\importance{Human microbiota is tightly associated to the health status of a person. Here we analyse the microbial composition of several subjects under different conditions, over a time span that ranges from days to months. Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas. This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system.}

@@

%===================================== SECTIONS =======================================
%%% Introduction
\section*{Introduction}
\input{mSys_intro_rev.tex}
%\clearpage
%%% Result
\section*{Results}
\input{mSys_results_rev.tex}
%\clearpage
%%% Discussion
\section*{Discussion}
\input{mSys_discussion_rev.tex}
%\clearpage    
%%% Material & Methods
\section*{Materials and Methods}
\input{mSys_MandM_rev.tex}   
%\clearpage

@@

%% Discussion
\bibitem {kilpatrick} {\bf Kilpatrick a M, Ives a R.} 2003. Species interactions can explain Taylor’s power law for ecological time series. Nature {\bf 422}:65–68.
\bibitem {ballantyne} {\bf Ballantyne IV F, J. Kerkhoff A.} 2007. The observed range for temporal mean-variance scaling exponents can be explained by reproductive correlation. Oikos {\bf 116}:174–180.
\bibitem {joao} {\bf Stein RR, Bucci V, Toussaint NC, Buffie CG, Rätsch G, Pamer EG, Sander C, Xavier JB.}  2013. Ecological modeling from time-series inference: insight into dynamics and stability of intestinal microbiota. PLoS Comput Biol {\bf 9}:e1003388.
\bibitem {mehta} {\bf Fisher CK, Mehta P.} 2014. Identifying keystone species in the human gut microbiome from metagenomic timeseries using sparse linear regression. PLoS One {\bf 9}:e102451.
\bibitem {bucci}  {\bf Bucci V, Tzen B, Li N, Simmons M, Tanoue T, Bogart E, Deng L, Yeliseyev V, Delaney ML, Liu Q, Olle B, Stein RR, Honda K, Bry L, Gerber GK.} 2016. MDSINE: Microbial Dynamical Systems INference Engine for microbiome time-series analyses. Genome Biol {\bf 17}:121.
\bibitem {koenig} {\bf Koenig JE, Spor A, Scalfone N, Fricker AD, Stombaugh J, Knight R, Angenent LT, Ley RE.} 2011. Succession of microbial consortia in the developing infant gut microbiome. Proc Natl Acad Sci {\bf 108}:4578–4585.
\bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
\bibitem {sislands1} {\bf Ng SC, Lam EFC, Lam TTY, Chan Y, Law W, Tse PCH, Kamm M a, Sung JJY, Chan FKL, Wu JCY.} 2013. Effect of probiotic bacteria on the intestinal microbiota in irritable bowel syndrome. J Gastroenterol Hepatol {\bf 28}:1624–31.
\bibitem {sislands2} {\bf Carroll IM, Ringel-Kulka T, Siddle JP, Ringel Y.} 2012. Alterations in composition and diversity of the intestinal microbiota in patients with diarrhea-predominant irritable bowel syndrome. Neurogastroenterol Motil {\bf 24}:521-30
\bibitem {sislands3} {\bf Xu D, Gao J, Gillilland M, Wu X, Song I, Kao JY, Owyang C.} 2014. Rifaximin alters intestinal bacteria and prevents stress-induced gut inflammation and visceral hyperalgesia in rats. Gastroenterology {\bf 146}:484–496.e4.
\bibitem {sislands4} {\bf Rooks MG, Veiga P, Wardwell-scott LH, Tickle T, Segata N, Michaud M, Gallini CA, Beal C, van Hylckama-Vlieg JET, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS, Hylckama-vlieg JET Van, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS.} 2014. Gut microbiome composition and function in experimental colitis during active disease and treatment-induced remission. ISME J {\bf 8}:1403–17.
\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.

### 2016-12-21 mSystems/rev/mSys_MandM_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

%%
%\small
\subsection*{Model} \label{sec:model}

We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
\begin{linenomath}
\begin{equation*}
\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
\end{equation*}
\end{linenomath}
where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, . The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
\begin{linenomath}
\begin{equation*}
\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
\end{equation*}
\end{linenomath}
The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V. For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V. The steady-state solution of the Fokker-Planck equation is given by:
\begin{eqnarray}
P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest. The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$). For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$). The phase transition region can be calculated analytically in particular cases:
\begin{eqnarray}
F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
In many physical systems (Brownian motion is the classical example \cite{Einstein}), the two terms of the Langevin equation are related.  The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}. The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations 
of the microbiota, assuming that it is in equilibrium (the ordered phase).

@@

The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}. Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold. Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure \ref{fig:taxlev1}).

@@

%%
\subsubsection*{Taxa level robustness}
We selected genus as taxonomic level for the subsequent steps of our work. In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see last subsection of Material and Methods), we tested two different data sets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. The former row of subfigures is for 16S, where levels are family (blue circles) vs. genus (purple triangles). Whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs. species (purple triangles). The left column shows the raw results and the right column plots the standardized results (see Standardization under Material and Methods).}
\label{fig:taxlev1}
\end{figure}

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{fig:taxlev2}
\end{figure}

@@

%\normalsize

>>>>

%%
%TC:ignore
%\small
\subsection*{Model} \label{sec:model}

We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
\begin{linenomath}
\begin{equation*}
\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
\end{equation*}
\end{linenomath}
where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, . The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
\begin{linenomath}
\begin{equation*}
\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
\end{equation*}
\end{linenomath}
The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V. For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V. The steady-state solution of the Fokker-Planck equation is given by:
\begin{eqnarray*}
P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \\
P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
\end{eqnarray*}
where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest. The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$). For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$). The phase transition region can be calculated analytically in particular cases:
\begin{eqnarray*}
F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1,\\
F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray*}
where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
In many physical systems (Brownian motion is the classical example \cite{Einstein}), the two terms of the Langevin equation are related.  The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}. The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations 
of the microbiota, assuming that it is in equilibrium (the ordered phase).

@@

The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}. Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold. Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure \ref{supfig:taxlev1}).

@@

%%
\subsubsection*{Taxa level robustness}
We selected genus as taxonomic level for the subsequent steps of our work. In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see last subsection of Material and Methods), we tested two different data sets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Supplementary Figures \ref{supfig:taxlev1} (overview) and \ref{supfig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

%TC:endignore
\begin{supfig} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. The former row of subfigures is for 16S, where levels are family (blue circles) vs. genus (purple triangles). Whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs. species (purple triangles). The left column shows the raw results and the right column plots the standardized results (see Standardization under Material and Methods).}
\label{supfig:taxlev1}
\end{supfig}

\begin{supfig} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{supfig:taxlev2}
\end{supfig}
%TC:ignore

@@

%\normalsize
%TC:endignore
